Univalent Universes and Completness of Reversible Programs: \\ Featherweight HoTT

### 2017-03-02 upi.tex
starting to explain things carefully

% \usepackage{fdsymbol}
\usepackage{bbold}
\usepackage{bussproofs}
\usepackage{keystroke}
\usepackage{comment}

@@

>>>>

% \usepackage{fdsymbol}
\usepackage{bbold}
\usepackage{bussproofs}
\usepackage{keystroke}
\usepackage{comment}
\usepackage{tikz}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Preliminaries}

[Following the HoTT book and the slides on ``A Characterization of Univalent
Fibrations'' by Dan Christensen.]

We think of a type family $P : A \to \mathcal{U}$ as a \emph{fibration} with
base space $A$, with $P(x)$ being the fiber over $x$, and with
$\Sigma_{(x:A)} P(x)$ being the \emph{total space} of the fibration, and with
first projection $\Sigma_{(x:A)} P(x) \to A$. In the figure below, the pairs
$(x,u)$ and $(y,p_*(y))$ are in the total space. A critical property of
fibrations is that a path $p : x \equiv y$ in the base space $A$ and a point $u$
in the fiber over $x$ induce a function that \emph{lifts} the path to the total
space: the lifted path starts at $(x,u)$ and ends at $(y,p_*(u))$ for some point
$p_*(u)$ in the fiber over $y$ that depends on $p$ and $u$.

\begin{center}
\begin{tikzpicture}[scale=0.7,every node/.style={scale=0.7}]]
  \draw (-3,0) ellipse (1.5cm and 3cm);
  \draw (3,2) ellipse (0.5cm and 1cm);
  \draw (3,-2) ellipse (0.5cm and 1cm);
  \node[blue,ultra thick,above] at (-3,3) {$A$};
  \node[blue,ultra thick,above] at (3,3) {$P(x)$ (fiber over $x$)};
  \node[blue,ultra thick,below] at (3,-3) {$P(y)$ (fiber over $y$)};
  \draw[fill] (-3,1.5) circle [radius=0.025];
  \draw[fill] (-3,-1.5) circle [radius=0.025];
  \draw[left,cyan,thick] (-3,1.5) -- (-3,-1.5);
  \node[above] at (-3,1.5) {$x$};
  \node[below] at (-3,-1.5) {$y$};
  \draw[fill] (3,1.8) circle [radius=0.025];
  \draw[fill] (3,-1.8) circle [radius=0.025];
  \node[above] at (3,1.8) {$u$};
  \node[below] at (3,-1.8) {$p_*(u)$};
  \node[left,cyan] at (-3,0) {$p$};
  \draw[->,red,dashed,ultra thick] (-3,1.5) to [out=45, in=135] (2.4,2.5);
  \draw[->,red,dashed,ultra thick] (-3,-1.5) to [out=-45, in=-135] (2.4,-2.5);
\end{tikzpicture}
\end{center}
